### 2019-06-04 doc/chapters/introduction.tex
fixed footnotes

A powerful method to check whether a program works correctly is formal verification.
This is a systematic process based on mathematical modeling.
With formal verification all possible inputs are explored algorithmically and exhaustively.
This is in contrast to testing, where only predefined inputs are tested.
The correctness of a program is analyzed relative to its formal specification.
Formal specification is a mathematical description of a software behavior that can be used by the verification tool to verify the code.\cite{sanghavi:formal_verification}

@@

Stainless is developed by "Lab for Automated Reasoning and Analysis" (LARA) at EPFL's School of Computer and Communication Sciences.
Using this framework we can verify the correctness of Scala programs.
Stainless statically verifies that a program satisfies a given specification and that a program will not crash at runtime.
The framework explores all possible inputs and finds a counter examples for possible failures in a program which violate the given specification.\cite{Stainless:introduction}

@@

On invoking, Stainless tries to prove that the postcondition always holds, assuming the given precondition does hold.\cite{Stainless:introduction}

@@

Stainless can produce 3 outcomes of postcondition verification: \textit{valid}, \textit{invalid} and \textit{unknown}.
If the postcondition is \textit{valid} Stainless could prove that for any inputs constrained in the precondition, the postcondition always holds.
Reporting the postcondition as \textit{invalid} the framework could find at least one counterexample which satisfies the precondition but violates the postcondition.
If Stainless is unable to prove the postcondition or find a counterexample it reports the outcome \textit{unknown}.
In this case a timeout or an internal error occurred.
Furthermore, Stainless checks for calls in the code to the function with invalid arguments violating the precondition.
For example if we invoke factorial with a negative number it reports \textit{invalid}.\cite{Stainless:introduction}

@@

Stainless supports only a subset of Scala.
They call is Pure Scala.
You can find the specification of Pure Scala on the Stainless website.\cite{Stainless:pure_scala}

In addition, Stainless has its own library with annotations, reimplementaion of some core data types, collections and input-output functions and more.
Some of them are described on the website.\cite{Stainless:library}
There are more details about library in the source code of Stainless on GitHub in the folder \textit{frontends/library/stainless}.\cite{Stainless:github}

@@

In the work some fragments of the code of Bitcoin-S has to be verified.
Bitcoin-S is an open source Scala implementation of the Bitcoin protocol. 
There are different projects belonging to Bitcoin-S.
Protocol data structures are defined in the package \textbf{bitcoin-s-core}.
This package is the main target to be verified because it specifies the base of Bitcoin protocol, and it is crucial for correct protocol functionality.
There is also the package \textbf{bitcoind-rpc} which is an RPC client implementation to communicate with \textit{bitcoind}, daemon of Bitcoin Core running on a machine.
Developers of Bitcoin-S are working on the implementation of a lightweight bitcoin client in the package \textbf{bitcoin-s-spv-node}.
The whole list of packages can be found on GitHub \cite{BitcoinSProject}.

>>>>

A powerful method to check whether a program works correctly is formal verification.
This is a systematic process based on mathematical modeling.
With formal verification all possible inputs are explored algorithmically and exhaustively.
This is in contrast to testing, where only predefined inputs are tested.
The correctness of a program is analyzed relative to its formal specification.
Formal specification is a mathematical description of a software behavior that can be used by the verification tool to verify the code. \cite{sanghavi:formal_verification}

@@

Stainless is developed by "Lab for Automated Reasoning and Analysis" (LARA) at EPFL's School of Computer and Communication Sciences.
Using this framework we can verify the correctness of Scala programs.
Stainless statically verifies that a program satisfies a given specification and that a program will not crash at runtime.
The framework explores all possible inputs and finds a counter examples for possible failures in a program which violate the given specification. \cite{Stainless:introduction}

@@

On invoking, Stainless tries to prove that the postcondition always holds, assuming the given precondition does hold. \cite{Stainless:introduction}

@@

Stainless can produce 3 outcomes of postcondition verification: \textit{valid}, \textit{invalid} and \textit{unknown}.
If the postcondition is \textit{valid} Stainless could prove that for any inputs constrained in the precondition, the postcondition always holds.
Reporting the postcondition as \textit{invalid} the framework could find at least one counterexample which satisfies the precondition but violates the postcondition.
If Stainless is unable to prove the postcondition or find a counterexample it reports the outcome \textit{unknown}.
In this case a timeout or an internal error occurred.
Furthermore, Stainless checks for calls in the code to the function with invalid arguments violating the precondition.
For example if we invoke factorial with a negative number it reports \textit{invalid}. \cite{Stainless:introduction}

@@

Stainless supports only a subset of Scala.
They call is Pure Scala.
You can find the specification of Pure Scala on the Stainless website\footnote{\url{https://epfl-lara.github.io/stainless/purescala.html}}.

In addition, Stainless has its own library with annotations, reimplementaion of some core data types, collections and input-output functions and more.
Some of them are described on the website\footnote{\url{https://epfl-lara.github.io/stainless/library.html}}.
There are more details about library in the source code of Stainless on GitHub in the folder \textit{frontends/library/stainless}\footnote{\url{https://github.com/epfl-lara/stainless}}.

@@

In the work some fragments of the code of Bitcoin-S has to be verified.
Bitcoin-S is an open source Scala implementation of the Bitcoin protocol. 
There are different projects belonging to Bitcoin-S.
Protocol data structures are defined in the package \textbf{bitcoin-s-core}.
This package is the main target to be verified because it specifies the base of Bitcoin protocol, and it is crucial for correct protocol functionality.
There is also the package \textbf{bitcoind-rpc} which is an RPC client implementation to communicate with \textit{bitcoind}, daemon of Bitcoin Core running on a machine.
Developers of Bitcoin-S are working on the implementation of a lightweight bitcoin client in the package \textbf{bitcoin-s-spv-node}.
The whole list of packages can be found on GitHub\footnote{\url{https://github.com/bitcoin-s}}.

### 2019-06-04 doc/chapters/towards_verifying_addition.tex
fixed footnotes

This is, due to the internal design of Scala and Java.
It's possible to reason about case object but not about object.
This needs a fundamental knowledge of Scala and some functional paradigms that should not be part of this thesis.
The issue \mypound520 on Stainless GitHub gives some thoughts if you want to know more.

@@

This is a missing feature in Stainless.
It does not support upper type bound on the class itself.
To track this, there is the issue \mypound519 on GitHub.
\begin{lstlisting}[language=scala]
  sealed abstract class Number[T <: Number[T]] extends BasicArithmetic[T]  
\end{lstlisting}

@@

This was a missing feature of Stainless that was fixed on Mai 28, 2019 with pull request \mypound470.

>>>>

This is, due to the internal design of Scala and Java.
It's possible to reason about case object but not about object.
This needs a fundamental knowledge of Scala and some functional paradigms that should not be part of this thesis.
The issue \mypound520 on Stainless GitHub\footnote{\url{https://github.com/epfl-lara/stainless/issues/520}} gives some thoughts if you want to know more.

@@

This is a missing feature in Stainless.
It does not support upper type bound on the class itself.
To track this, there is the issue \mypound519 on GitHub\footnote{\url{https://github.com/epfl-lara/stainless/issues/519}}.
\begin{lstlisting}[language=scala]
  sealed abstract class Number[T <: Number[T]] extends BasicArithmetic[T]  
\end{lstlisting}

@@

This was a missing feature of Stainless that was fixed on Mai 28, 2019 with pull request \mypound470 on GitHub\footnote{\url{https://github.com/epfl-lara/stainless/pull/470}}.
